---
title: "Combined 2015 and 2023 Data Setup"
author: "Mark C. Otto"
eval: true
date: "5 October 2024"
format: docx
editor: source
---

## Introduction

We combine the 2015 and 2023 survey files: The spatial files are: `MEBoundrySF`, `StrSF`, `StrPlotSF`, `ColonySF`

The data files are `GRTSLoc` `tblFlightInfo` `tblPlotDetails` `tblObservations` `tblStaff`, `RecObs`.

Reference files: `StrArea`

These are combined over years. The flight, plot, and observation files combine into `RecObs`. Do I add the `begplot` and `endplot` in `r`RecObs\` when I can get that information from Danielle's tables? Why don't I use her tables and create the list, plot, and caphist files for information from them.

I follow Danielle D'Auria's convention of using `_PK` of a primary key variable and `_FK` for a foreign key variable.

The result can be an `SQLite3` database and a `R Savefile`.

```{r SetUp}
library(readr)
library(readxl)
library(sf)
library(dplyr)

library(stringr)
library(tidyr)
library(purrr)
library(ggplot2)
library(RSQLite)
source(here::here("R/RDB2Data.R"))

# Survey years
CurrentYr <- 2023
cYear <- c(2015, CurrentYr)
nYear <- length(cYear)

# Survey frames
cFrame <- c("List", "Area")
nFrame <- length(cFrame)

# Survey strata
cDFStr <- c("SC", "Co", "ND")
nStr <- length(cDFStr)

# Make the year-stratum levels
cYrStr <- apply(expand_grid(Year = cYear, Str = cDFStr), 1, function(x)
  paste(x[1], x[2], sep = "."))
nYrStr <- length(cYrStr)

# Make the year-frame-stratum levels.
# Should call cYFS.
cFS <- apply(expand_grid(Year = cYear, Frame = cFrame, Str = cDFStr), 1, function(x)
  paste(x[1], x[2], x[3], sep = "."))
nFS <- length(cFS)

lWrite <- FALSE

```

## CRS Projections

The spatial projections or CRS codes do not seem to define the projections correctly, so save them: Alpers equal area, UTS, and WGS84 (World Geodetic System 1984).

```{r CRSProjections}
CRS.WGS84 <- st_crs("EPSG:4326")
tStrPlotSF <- st_read("~/Profesional/GBHE.2023/GIS/StrPlotSp.shp")
CRS.Alpers <- st_crs(tStrPlotSF)
tColonySp <- st_read("~/Profesional/GBHE.2023/GIS/2015/GBHE_ColonyPlotJoin.shp")
CRS.UTM19N <- st_crs(tColonySp)
if (lWrite) {
  save(CRS.WGS84, file = here::here("data/CRS.WGS84.RData"))
  save(CRS.Alpers, file = here::here("data/CRS.Alpers.RData"))
  save(CRS.UTM19N, file = here::here("data/CRS.UTM19N.RData"))
}

```

## Maine State Boundary Layer

The state boundary polygon, [Maine State Boundary Polygon Feature](https://maine.hub.arcgis.com/datasets/maine::maine-state-boundary-polygon-feature/explore), is `WGS 84`. This is just a reference layer, so leave the variables as is. Transform it to Alpers and save as a KML file

```{r MEBoundary}
tMEBoundarySF <- st_read("~/Profesional/GBHE.2023/GIS/Maine_State_Boundary_Polygon_Feature.shp")
MEBoundarySF <- st_transform(x = tMEBoundarySF, crs = CRS.Alpers)
if (lWrite) {
  st_write(
    MEBoundarySF,
    dsn = here::here("gis/MEBoundarySF.kml"),
    delete_layer = TRUE
  )
}
rm(tMEBoundarySF)

```

## Stratum Layer

Stratum, `StrSF`. Don't worry about the complication of list and area strata that are not involved in this survey.

Maine is supposed to have 91,646km². Danielle may have removed some water area. Use the areas in `StrArea`.

```{r Stratum}
StrArea <- RDB2Data(Tbl = "StrArea",
                    RowNames = "DFStr",
                    RDB = "~/Profesional/GBHE.2023/extdata/2023/DFS2023.sq3") %>%
  select(Str_PK = DFStr, cStratum, KMSq)
fStr <- factor(StrArea$Str_PK, StrArea$Str_PK)

tStrSF <- st_read("~/Profesional/GBHE.2023/GIS/GBHE_Stratum.shp") %>%
  mutate(KMSq = AREA / 10 ^ 6, cStratum = NAME2_) %>%
  select(cStratum, KMSq)

# `dplyr` joins removes the `sf` class.
tStrSF2 <- st_as_sf(inner_join(tibble::as_tibble(
  x = list(Str_FK = fStr, cStratum = StrArea$cStratum)
), tStrSF, by = "cStratum") %>%
  arrange(Str_FK))

StrSF <- st_transform(x = tStrSF2, crs = CRS.Alpers)
if (lWrite) {
  st_write(StrSF,
           dsn = here::here("gis/StrSF.kml"),
           delete_layer = TRUE)
}
rm(tStrSF, tStrSF2)

```

## Colony List for both years

We should start with the colony list we had in 2015, add the new colonies found in the 2015 survey and any in the following eight years.

What is odd is ME.0712 was dropped from the list and we had new colonies at least found in the survey. If ME.0712 was in active since the last survey, then there is good reason to drop it.

The `AnnObs` table is a good way to create the list colonies for each survey, but we don't have a current set of those observations.

To add the 2015 new colonies, use the locations and match them to the `AllColonies` list. Ask Danielle.

Make the spatial layer out of the `.csv` files.

```{r ListColony}
tColony2015 <- RDB2Data(Tbl = "Nest",
                        RowNames = "FWSNestID",
                        RDB = "~/Profesional/GBHE.2023/extdata/2015/DFS2015.sq3") %>%
  select(
    FWSColonyID_PK = FWSNestID,
    Str_FK = LstStr,
    PlotID_FK = PlotID,
    Lng,
    Lat,
    LastYr,
    Status,
    nNest,
    nActive,
    Comment
  )

# These are all the colonies.  445 of them so select from the last list
# Where are the new colonies found in the last survey.
# Is in UTM 19N
RC2XY <- read_csv("~/Profesional/GBHE.2015/ExtData/RC2XY.csv", col_types = "c-ii---")
tCol2023SF <- inner_join(
  st_read("~/Profesional/GBHE.2023/GIS/2023/ListFrameNewPlots.shp") %>%
    mutate(
      FWSColonyID_PK = sprintf("ME.%04d", COLONYNO),
      Str_FK = c(
        "South Central" = "SC",
        "Coastal" = "Co",
        "North and Downeast" = "ND"
      )[NAME2_],
      nNest = AvgMaxNest
    ),
  RC2XY,
  by = c("Col", "Row")
) %>%
  select(FWSColonyID_PK, Str_FK, PlotID_FK = PlotID, nNest)

# Add the lat-longs
LatLongSF <- st_transform(x = tCol2023SF, crs = CRS.WGS84)
LatLong <- st_drop_geometry(LatLongSF %>%
                              mutate(Lng = unlist(map(
                                LatLongSF$geometry, 1
                              )), Lat = unlist(map(
                                LatLongSF$geometry, 2
                              ))) %>%
                              select(Lng, Lat))

# Add the long-lat to the nestlist
Col2023SF <- bind_cols(tCol2023SF, LatLong) %>%
  mutate(
    LastYr = NA,
    Status = NA,
    nActive = NA,
    Comment = NA
  ) %>%
  select(FWSColonyID_PK,
         Str_FK,
         PlotID_FK,
         Lng,
         Lat,
         LastYr,
         Status,
         nNest,
         nActive,
         Comment)

tColony2015SF <- inner_join(
  st_read("~/Profesional/GBHE.2023/GIS/2015/GBHE_ColonyPlotJoin.shp") %>%
    mutate(FWSColonyID_PK = sprintf("ME.%04d", COLONYNO)) %>%
    select(FWSColonyID_PK),
  tColony2015,
  by = "FWSColonyID_PK"
)

# For now, make an entry for each year.  We should get these from
# and annual observations file like `AnnObs`.
ColonySF <- st_as_sf(bind_rows(
  tColony2015SF %>%
    mutate(Year = 2015),
  Col2023SF %>%
    mutate(Year = 2023)
))  %>%
  select(
    FWSColonyID_PK,
    Year,
    Str_FK,
    PlotID_FK,
    Lng,
    Lat,
    LastYr,
    Status,
    nNest,
    nActive,
    Comment
  )

if (lWrite) {
  st_write(ColonySF,
           dsn = here::here("gis/ColonySF.kml"),
           delete_layer = TRUE)
}

rm(tColony2015, tColony2023, tColony2015SF)

```
## Non-Spatial Colony List

```{r NonSpatialColonies}
Colony <-  st_drop_geometry(ColonySF) %>%
  select(
    Year,
    FWSColonyID = FWSColonyID_PK,
    Str = Str_FK,
    PlotID = PlotID_FK,
    Lng,
    Lat,
    nNest,
    nActive
  )
if (lWrite) {
  readr::write_csv(Colony, here::here("data/Colony.csv"), na = "")
}

```
## Stratum-Plot Layer

I intersected or spatially joined the stratum with the plot layers, so use `StrPlotSF`. Later create the lat-long using `st_centroid()`. Also copy over the `Plot` table that is non spatial but has number of colonies and km^2^. The row and col come from `RC2XY` made in the Bald Eagle national grid.

```{r StrPlot}
tStrPlotSF <- st_read("~/Profesional/GBHE.2023/GIS/StrPlotSp.shp") %>%
  select(Str_FK = Str, PlotID_FK = PlotID, KMSq, Col, Row)

tLatLong <- st_transform(x = st_centroid(tStrPlotSF), crs = CRS.WGS84)
LatLong <- st_drop_geometry(tLatLong %>%
                              mutate(Lng = unlist(map(
                                tLatLong$geometry, 1
                              )), Lat = unlist(map(
                                tLatLong$geometry, 2
                              ))) %>%
                              select(Lng, Lat))

# Add the long-lat to the nestlist
StrPlotSF <- bind_cols(tStrPlotSF, LatLong) %>%
  select(Str_FK, PlotID_FK, KMSq, Lng, Lat, Col, Row)
if(lWrite) {
  st_write(StrPlotSF, here::here("GIS/StrPlotSF.kml"), delete_layer = TRUE)
}
rm(tStrPlotSF, tLatLong, LatLong)

```

## All Known Colonies

We need allcolonies to incorporate the 2015 new nests into the 2023 list and some colonies not in the original list were observered.

```{r AllColonies}
# NestList is the colony list.  Need to remove all the colonies without
# locations.
tColonySF <- read_csv(file = "~/Profesional/GBHE.2023/extdata/2023/COLONIES.csv", col_types = "ic--c---------cccc---dd---c-d------ii") %>%
  filter(!is.na(X_UTMNAD83)) %>%
  mutate(
    FWSColonyID = sprintf("ME.%04d", COLONYNO),
    Year = ifelse(Found2009, "2009", "<2009"),
    Comment = paste(
      SiteID,
      ColName,
      HabitatDesc,
      Directions,
      Comments,
      History,
      sep = ";"
    )
  ) %>%
  select(FWSColonyID, Nest = COLSIZE, Year, Comment, X_UTMNAD83, Y_UTMNAD83) %>%
  # 0919 was a duplicate.
  distinct() %>%
  st_as_sf(coords = c("X_UTMNAD83", "Y_UTMNAD83"),
           crs = CRS.UTM19N)

tLatLong <- st_transform(x = tColonySF, crs = CRS.WGS84)

LatLong <- st_drop_geometry(tLatLong %>%
                              mutate(Lng = unlist(map(
                                tLatLong$geometry, 1
                              )), Lat = unlist(map(
                                tLatLong$geometry, 2
                              ))) %>%
                              select(Lng, Lat))

# Add the long-lat to the nestlist
tColonySF2 <- bind_cols(tColonySF, LatLong)

AllColonySF <- st_intersection(
  StrPlotSF %>%
    select(Str = Str_FK, PlotID = PlotID_FK),
  st_transform(x = tColonySF2, crs = st_crs(StrPlotSF))
) %>%
  select(FWSColonyID, Year, Str, PlotID, Nest, Lng, Lat, Comment)

# Colony has more colonies than AllColonies
PickOne <- function(x){
  idx <- !is.na(x)
  return(x[idx][1])
}

tAllCol <- bind_rows(
  st_transform(AllColonySF, crs = CRS.UTM19N),
  #  st_drop_geometry(ColonySF) %>%
  st_transform(ColonySF, crs = CRS.UTM19N) %>%
    mutate(
      FWSColonyID = FWSColonyID_PK,
      Str = Str_FK,
      PlotID = PlotID_FK,
    ) %>%
    group_by(FWSColonyID, Year, Str, PlotID) %>%
    summarize(
      Year = min(Year, 2009),
      Nest = max(nNest, 0, na.rm = TRUE),
      Lng = mean(Lng, na.rm = TRUE),
      Lat = mean(Lat, na.rm = TRUE),
      Comment = PickOne(Comment),
      .groups = "drop"
    ) %>%
    mutate(Year = as.character(Year), ) %>%
    dplyr::select(FWSColonyID, Year, Str, PlotID, Nest, Lng, Lat, Comment)
) %>%
  distinct() %>%
  group_by(FWSColonyID, Year, Str, PlotID) %>%
  summarize(
    Year = min(Year, 2009),
    Nest = as.integer(max(Nest, 0, na.rm = TRUE)),
    # Del = "",
    Lng = mean(Lng, na.rm = TRUE),
    Lat = mean(Lat, na.rm = TRUE),
    Comment = PickOne(Comment),
    .groups = "drop"
  ) %>%
  arrange(FWSColonyID, desc(Nest)) %>%
  filter(!(duplicated(FWSColonyID)
           #&(is.na(Comment)|grepl("count", Comment))
           ))
AllColonySF <- tAllCol           

if (lWrite) {
  write_csv(st_drop_geometry(AllColonySF), here::here("data/AllColony.csv"), na = "")
  st_write(AllColonySF,
           here::here("gis/AllColonySF.kml"),
           delete_layer = TRUE)
}
rm(tColonySF, tColonySF2, tLatLong, LatLong)

```

## List Plot List

Doesn't this depend on the year the sample was taken as to what is put in the list? For now we use all colonies.

```{r List}
List <- st_drop_geometry(ColonySF) %>%
  select(Str = Str_FK,
         PlotID = PlotID_FK,
         nNest,
         # Should standardize: nActive or nAct
         nAct = nActive) %>%
  group_by(Str, PlotID) %>%
  summarize(
    nCol = n(),
    nNest = sum(nNest, na.rm = TRUE),
    nAct = sum(nAct, na.rm = TRUE),
    .groups = "drop"
  )
if (lWrite) {
  readr::write_csv(List, here::here("data/List.csv"), na = "")
}

```

## Identify the colonies used for known colonies.

```{r IntoBaldEagleForm}
#| eval: false
tColUsed <- read_csv("~/Profesional/GBHE.2023/extdata/2023/GBHE_ColonyPlotJoin.csv",
                     col_types = "---idd------ii----------------c-") %>%
  mutate(FWSColonyID = sprintf("ME.%04d", COLONYNO),
         DatEntered = "2022-01-01") %>%
  select(FWSColonyID, DatEntered)

# # Identfy the colonies used by 2022 for the `DatEntered` and 2008 otherwise.
# ColoniesSF <- left_join(tColonySF2, tColUsed, by = "FWSColonyID") %>%
#   mutate(
#     St = "ME",
#     StNestID = as.integer(sub("ME.", "", FWSColonyID)),
#     LocAcc = .1,
#     Location = "",
#     Access = "",
#     Owner = "Danielle D'Auria",
#     ListYr = 2008,
#     DatEntered = ifelse(is.na(DatEntered), "2008-01-01", DatEntered),
#     Comment = ""
#   ) %>%
#   select(
#     FWSColonyID  ,
#     St       ,
#     StNestID ,
#     PlotID   ,
#     Str      ,
#     Lng      ,
#     Lat      ,
#     LocAcc   ,
#     Location ,
#     Access   ,
#     Owner    ,
#     ListYr   ,
#     DatEntered,
#     Comment
#   ) %>%
#   arrange(FWSColonyID)
# if (lWrite) {
#   st_write(ColonySF, here::here("GIS/AllColonySF.kml"), delete_layer = TRUE)
# }

```

## Plot Non-Spatial

Add the nests too.

```{r Plot}
tPlot <- st_drop_geometry(StrPlotSF) %>%
  select(Str = Str_FK, PlotID = PlotID_FK, KMSq, Lng, Lat)

Plot <- left_join(tPlot, List, by = c("Str", "PlotID")) %>%
  mutate(
    nCol = ifelse(is.na(nCol), 0, nCol),
    nNest = ifelse(is.na(nNest), 0, nNest),
    nAct = ifelse(is.na(nAct), 0, nAct)
  ) %>%
  select(Str, PlotID, Lng, Lat, KMSq, nCol, nNest, nAct)
if (lWrite) {
  write_csv(Plot, here::here("GIS/Plot.csv"), na = "")
}
rm(tPlot)

```

## GRTS Sample

Combine the current year GRTS sample file with the past years. Add the year onto the tables.  Archive `GRTSLoc` that has all the information used for the sample design.  Make `GRTSSmp` that identifies the samples flown and keep that for the estimation.

L79 was sampled, so correct `EvalStatus`.

```{r GetFromDB}
tGRTSLoc <- RDB2Data(
  Tbl = "GRTSLoc",
  RowNames = "siteID",
  RDB = "~/Profesional/GBHE.2023/extdata/2015/DFS2015.sq3",
  envir = NULL
) %>%
  mutate(Year = 2015)

tGRTSCurrent <- RDB2Data(
  Tbl = "GRTSLoc",
  RowNames = "siteID",
  RDB = "~/Profesional/GBHE.2023/extdata/2023/DFS2023.sq3",
  envir = NULL
) %>%
  mutate(Year = CurrentYr)


tGRTSLoc2 <- bind_rows(tGRTSLoc, tGRTSCurrent) %>%
  mutate(
    Str = stratum,
    siteID = Plot,
    EvalStatus = ifelse(is.na(EvalStatus), FALSE, EvalStatus == "Evaluated"),
    EvalStatus = ifelse(siteID == "L79", TRUE, EvalStatus)
  )

GRTSLoc <- inner_join(tGRTSLoc2,
                      Plot  %>%
                        select(Str, PlotID, KMSq, nAct),
                      by = c("Str", "PlotID")) %>%
  select(
    Year,
    siteID,
    Str,
    PlotID,
    SmpType,
    xcoord,
    ycoord,
    mdcaty,
    wgt,
    KMSq,
    nAct,
    panel,
    EvalStatus,
    EvalReason,
    Comment
  )
if (lWrite) {
  readr::write_csv(GRTSLoc, here::here("data/GRTSLoc.csv"), na = "")
}
rm(tGRTSLoc, tGRTSLoc2, tGRTSCurrent)

```
## `GRTSSmp` for the Estimation

```{r GRTSSmp}
GRTSSmp <- GRTSLoc %>%
  filter(EvalStatus) %>%
  select(Year, siteID, Str, PlotID, SmpType, KMSq, nAct)
if (lWrite) {
  readr::write_csv(GRTSSmp, here::here("data/GRTSSmp.csv"), na = "")
}


```

The colony observations, `RecObs` are created from `tblFlightInfo`, `tblObservations`, `tblPlotDetails`, and `tblStaff`.

First combine the flight and plot files. Problems with commas in the weather notes, `WxNotes` in the 2023 table. Put quotes around the field.

```{r FlightPlot}
tblFlightInfo2015 <- readr::read_csv("~/Profesional/GBHE.2015/ExtData/GBHEDualFrame2/tblFlightInfo.csv",
                                     col_types = "iiiiccccccc")

tblFlightInfo2023 <- read_csv("~/Profesional/GBHE.2023/extdata/2023/tblFlightInfo.csv",
                              col_types = "iiiiccccccc")
#  %>%mutate(Start = as.character(Start), End = as.character(End))
tblFlightInfo <- bind_rows(tblFlightInfo2015, tblFlightInfo2023) %>%
  mutate(
    Date = sprintf("%04d-%02d-%02d", DateYear, DateMonth, DateDay),
    Start = as.POSIXct(Start, format = "%H:%M:%S"),
    End = as.POSIXct(End, format = "%H:%M:%S"),
    Comment = paste(ifelse(is.na(WxNotes), "", WxNotes), ifelse(is.na(Comments), "", Comments), sep = "; ")
  ) %>%
  select(
    Flight_PK,
    Year = DateYear,
    Date,
    Start,
    End,
    Pilot,
    Front = FObs,
    Rear = RObs,
    Comment
  )
if (lWrite) {
  readr::write_csv(tblFlightInfo, here::here("data/tblFlightInfo.csv"), na = "")
}
rm(tblFlightInfo2015, tblFlightInfo2023)

```

## Merge the plot tables

Combine with `GRTSLoc` to add `Str` and `PlotID`. The `OOS` will not have stratum or plot ID.

```{r PlotDetails}
tblPlotDetails2015 <- readr::read_csv("~/Profesional/GBHE.2015/ExtData/GBHEDualFrame2/tblPlotDetails.csv",
                                      col_types = "iclccccci") %>%
  mutate(Year = 2015)

tblPlotDetails2023 <- read_csv("~/Profesional/GBHE.2023/extdata/2023/tblPlotDetails.csv",
                               col_types = "iiclccccc") %>%
  mutate(Year = CurrentYr)

ttblPlotDetails <- bind_rows(tblPlotDetails2015, tblPlotDetails2023) %>%
  mutate(
    BegArea = as.POSIXct(BegArea, format = "%H:%M:%S"),
    EndArea = as.POSIXct(EndArea, format = "%H:%M:%S"),
    BegList = as.POSIXct(BegList, format = "%H:%M:%S"),
    EndList = as.POSIXct(EndList, format = "%H:%M:%S"),
    Comment = ifelse(`Combined?`, paste("Combined.", ifelse(
      is.na(Comments), "", Comments
    )), Comments)
  )

tblPlotDetails <-
  left_join(
    ttblPlotDetails,
    GRTSLoc %>%
      select(Year, SiteID = siteID, Str, PlotID),
    by = c("Year", "SiteID")
  ) %>%
  select(
    PlotDetails_PK,
    Year,
    Str,
    Flight_FK,
    SiteID,
    PlotID,
    BegArea,
    EndArea,
    BegList,
    EndList,
    Comment
  )
if (lWrite) {
  readr::write_csv(tblPlotDetails, here::here("data/tblPlotDetails.csv"), na = "")
}
rm(tblPlotDetails2015, tblPlotDetails2023, ttblPlotDetails)

```

## 2015 Observations

I am assuming the coordinates are in UTM 19N. Convert them to lat-longs for the few that have coordinates. These are important since the GBHE ones are the new colonies and we have to add them to the 2023 colony list.

Save `LatLong` to correct `RecObs2015` too.

```{r 2015Observations}
ttblObservations2015 <- readr::read_csv(
  "~/Profesional/GBHE.2015/ExtData/GBHEDualFrame2/tblObservations.csv",
  col_types = "iclicciiiicccicii"
) %>%
  mutate(Year = 2015) %>%
  select(
    Obs_PK,
    PlotDetails_FK,
    WaypointID,
    `NewColony?`,
    WBC_ID,
    Strata,
    Time,
    CaptHist,
    FObsTot,
    FObsAct,
    RObsTot,
    RObsAct,
    Species,
    Habitat,
    Comments,
    Latitude,
    Longitude,
    Year
  )

# For only those observations with locations.
Loc2015SF <- ttblObservations2015 %>%
  filter(!is.na(Latitude)) %>%
  mutate(xUTM = Longitude, yUTM = Latitude) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = CRS.UTM19N) %>%
  st_transform(crs = CRS.WGS84)

LatLong <- st_drop_geometry(Loc2015SF %>%
                              mutate(Lng = unlist(map(
                                Loc2015SF$geometry, 1
                              )), Lat = unlist(map(
                                Loc2015SF$geometry, 2
                              ))) %>%
                              select(Obs_PK, xUTM, yUTM, Lat, Lng))

tblObservations2015 <- left_join(ttblObservations2015, LatLong, by = "Obs_PK") %>%
  select(
    Obs_PK,
    PlotDetails_FK,
    WaypointID,
    `NewColony?`,
    WBC_ID,
    Strata,
    Time,
    CaptHist,
    FObsTot,
    FObsAct,
    RObsTot,
    RObsAct,
    Species,
    Habitat,
    Comments,
    Latitude = Lat,
    Longitude = Lng,
    Year
  )

rm(ttblObservations2015, Loc2015SF)

```

## 2023 Observations

Make edits that I would have made on `RecObs`.

```{r 2023Observations}

tblObservations2023 <- read_csv("~/Profesional/GBHE.2023/extdata/2023/tblObservations.csv",
                                col_types = "iicliccciiiicccii") %>%
  mutate(Year = CurrentYr)

```

## Combine Observations

Need to edit the 2023 observations like Danielle and I did with `RecObs`. and add all the locations for the known colonies. Also, many of the nests are not on the colony list we used for 2023. 1. They don't have the 2015 new nests and 2. observed ones that are not on the list we used but are on the all colonies list.

To get this done, combine `RecObs`.

```{r Combine2015-2023}
tblObservations <- bind_rows(tblObservations2015, tblObservations2023)  %>%
  mutate(
    FWSColonyID = ifelse(is.na(WBC_ID), NA, ifelse(
      `NewColony?`,
      sprintf("new.%04d", WBC_ID),
      sprintf("ME.%04d", WBC_ID)
    )),
    Str = Strata,
    CapHist = tolower(CaptHist),
  ) %>%
  select(
    Year,
    Obs_PK,
    PlotDetails_FK,
    Lng = Longitude,
    Lat = Latitude,
    AOU = Species,
    CapHist,
    Location = Habitat,
    FWSColonyID,
    fNest = FObsTot,
    rNest = RObsTot,
    fAct = FObsAct,
    rAct = RObsAct,
    Comment = Comments
  )
if (lWrite) {
  readr::write_csv(tblObservations,
                   here::here("data/tblObservations.csv"),
                   na = "")
}

rm(tblObservations2015, tblObservations2023)

```

## List the new colonies, since in `tblObservations` and `RecObs` the we use the new colony id instead of a code with `new.`.

```{r NewColonies}
NewColonies <- readr::read_csv(here::here("data/NewColonies.csv"), col_types = "ic")

```
## Combine `RecObs` from `tbl*` data files

```{r CombineRecObs}
tRecObs <- inner_join(tblFlightInfo,
                      tblPlotDetails,
                      by = c(Flight_PK = "Flight_FK", "Year")) %>%
  mutate(Comment = paste(Comment.x, Comment.y, sep = ";")) %>%
  select(
    Flight_PK,
    PlotDetails_PK,
    Year,
    Date,
    BegFlight = Start,
    EndFlight = End,
    Pilot ,
    Front,
    Rear,
    Str ,
    siteID = SiteID,
    PlotID,
    BegArea,
    EndArea,
    BegList,
    EndList,
    Comment
  )

# I should do a spatial join for every colony observed in case there
# are any strata that split sample plots.
tRecObs2 <- left_join(
  tRecObs %>%
    select(
      Flight_PK,
      PlotDetails_PK,
      Year,
      siteID,
      PlotID,
      Date,
      Pilot,
      Front,
      Rear,
      Comment
    ),
  GRTSLoc %>%
    select(Year, siteID, PlotID, Str),
  by = c("Year", "siteID", "PlotID")
) %>%
  # Should not need this
  distinct()

# Merge the flight, plot, and observation tables to make `RecObs`.
# Keep all the flight, plot, and observation keys for reference.
# All the sampled colonies have a stratum and plot.
# Don't match the bald eagle RecObs.  Obs_PK is equivalent to OBJECTID.
tRecObs3 <- inner_join(tRecObs2,
                       tblObservations,
                       by = c(PlotDetails_PK = "PlotDetails_FK", "Year")) %>%
  mutate(
    FWSColonyID = sub("new", "ME", FWSColonyID),
    Comment = paste(Comment.x, Comment.y, sep = ";")
  ) %>%
  select(
    Obs_PK,
    Year,
    Str,
    siteID,
    PlotID,
    DatTim = Date,
    Pilot,
    Front,
    Rear,
    Lng,
    Lat,
    AOU,
    CapHist,
    Location,
    FWSColonyID,
    fNest,
    rNest,
    fAct,
    rAct,
    Comment,
    Flight_PK,
    PlotDetails_PK
  ) %>%
  distinct()

# Obs_PK = 151 was duplicated.  Find why.
tRecObs4 <- left_join(
  tRecObs3,
  AllColonySF %>%
    select(
      FWSColonyID,
      tStr = Str,
      tPlotID = PlotID,
      Lng,
      Lat #,
      # Str = Str_FK,
      # PlotID = PlotID_FK
    ),
  by = "FWSColonyID"
)  %>%
  mutate(
    Str = ifelse(is.na(Str), tStr, Str),
    PlotID = ifelse(is.na(PlotID), tPlotID, PlotID),
    Lng = ifelse(is.na(Lng.x), Lng.y, Lng.x),
    Lat = ifelse(is.na(Lat.x), Lat.y, Lat.x) #,
    # Str = ifelse(is.na(Str.x), Str.y, ifelse(
    #   Str.x == Str.y, Str.x, paste(Str.x, Str.y, sep = "!=")
    # )),
    # PlotID = ifelse(
    #   is.na(PlotID.x),
    #   PlotID.y,
    #   ifelse(
    #     PlotID.x == PlotID.y,
    #     PlotID.x,
    #     paste(PlotID.x, PlotID.y, sep = "!=")
    #   )
    # )
  )  %>%
  select(
    Obs_PK,
    Year,
    Str,
    siteID,
    PlotID,
    DatTim,
    Pilot,
    Front,
    Rear,
    Lng,
    Lat,
    AOU,
    CapHist,
    Location,
    FWSColonyID,
    fNest,
    rNest,
    fAct,
    rAct,
    Comment,
    Flight_PK,
    PlotDetails_PK
  ) %>%
  arrange(Obs_PK)

RecObs <- st_drop_geometry(tRecObs4)


if (lWrite) {
  readr::write_csv(RecObs, here::here("data/RecObs.MCO.csv"), na = "")
}

rm(tRecObs, tRecObs2, tRecObs3)

```

## Combine `RecObs` from year-specific `RecObs`

Use `Obs_FK` rather than `OBJECTID`.\
In 2015, change the UTMs to lat-longs. Add the locations for 2015 using the colony list.

```{r CombineRecObsYS}
#| eval: false
RecObs2015 <- RDB2Data(Tbl = "RecObs",
                       RowNames = "OBJECTID",
                       RDB = "~/Profesional/GBHE.2023/extdata/2015/DFS2015.sq3") %>%
  mutate(Yr = 2015,
         siteID = ifelse(siteID == "outplot", "OOS", siteID)) %>%
  mutate(FWSColonyID = FWSNestID)

RecObs2023 <- RDB2Data(Tbl = "RecObs",
                       RowNames = "OBJECTID",
                       RDB = "~/Profesional/GBHE.2023/extdata/2023/DFS2023.sq3") %>%
  mutate(FWSColonyID = stringr::str_replace(FWSNestID, "new\\.[:alpha:][:alpha:]", "ME"))

tRecObs <- bind_rows(RecObs2015, RecObs2023) %>%
  select(
    OBJECTID,
    Year = Yr,
    St,
    Svy,
    DatTim,
    Pilot,
    Front,
    Rear,
    Str,
    PlotID,
    siteID,
    Lng,
    Lat,
    AOU,
    CapHist,
    Location,
    FWSColonyID,
    fNest,
    rNest,
    fAct,
    rAct,
    Comment,
    Edit,
    File,
    Rec,
    FWSNestID
  )

tRecObs2 <- left_join(tRecObs,
                      AllColonySF %>% select(FWSColonyID, Lng, Lat, Str, PlotID),
                      by = "FWSColonyID")  %>%
  mutate(
    Lng = ifelse(is.na(Lng.x), Lng.y, Lng.x / 10 ^ 4),
    Lat = ifelse(is.na(Lat.x), Lat.y, Lat.x / 10 ^ 5),
    Str = ifelse(is.na(Str.x), Str.y, ifelse(
      Str.x == Str.y, Str.x, paste(Str.x, Str.y, sep = "!=")
    )),
    PlotID = ifelse(
      is.na(PlotID.x),
      PlotID.y,
      ifelse(
        PlotID.x == PlotID.y,
        PlotID.x,
        paste(PlotID.x, PlotID.y, sep = "!=")
      )
    )
  )  %>%
  select(
    OBJECTID,
    Year,
    St,
    Svy,
    DatTim,
    Pilot,
    Front,
    Rear,
    Str,
    PlotID,
    siteID,
    Lng,
    Lat,
    AOU,
    CapHist,
    Location,
    FWSColonyID = FWSNestID,
    fNest,
    rNest,
    fAct,
    rAct,
    Comment,
    Edit,
    File,
    Rec
  )

RecObsYS <- tRecObs2

```

## Edit 2023 Observations

```{r EditRecObs}
#| eval: false
# MCO:  The ME.0647 seems to be off c6.1v vs. c4.1r (2 over and 3 up).
# DDA:  This should be c6.1v.
#  This is probably due to the original data entry error where it was entered as 105.
# MCO:  ColList:  ME.0647 c6.1v SC -70.78627 43.17239
#   It was in a list plot, L63, with no nests.
#   c6.1v is not a sample plot, but this was the colony that was sampled?
#   ColList:  ME.0105 c4.1r SC -70.82679 43.61651
#   105 is in the c4.1r or the L63 sample plot.  Is 105 647?
# RecObs[25,"FWSColonyID"]
iObs <- which(RecObs$FWSColonyID == "ME.0647")
RecObs[iObs, "PlotID"] <- "c6.1v"
RecObs[iObs, "siteID"] <- "OOS"
RecObs[iObs, "Edit"] <- "Changed PlotID from c4.1r and siteID from L63."

# MCO:  new.SC.0917 may be just over the border, so go with what you marked it as.
# DDA:  This should be ci.18.
# MCO:  So it is outside the A16 ci.19 sample plot?
#   That would make it OOS.  I will change it.
#  RecObs[81,"FWSColonyID"]
iObs <- which(RecObs$FWSColonyID == "new.SC.0917")
RecObs[iObs, "PlotID"] <- "ci.18"
RecObs[iObs, "siteID"] <- "OOS"
RecObs[iObs, "Edit"] <- "Changed PlotID from ci.19 and siteID from A18."

# MCO: ch.1c is a split plot, so go with the Co part of the plot?
# DDA:   Yes, this colony (866) is in Co stratum.
# MCO: Good
#  RecObs[184,"FWSColonyID"]
iObs <- which(RecObs$FWSColonyID == "ME.0866")
RecObs[iObs, "Str"] <- "Co"
RecObs[iObs, "Edit"] <- "Changed Str from SC because it was a plot split between strata."

# MCO: ME.0721 is just over in the next plot, so should we stick with the
# assigned plot as it is the sample plot?
# DDA:  This should be c4.16
# MCO: c4.16 is not a sample plot, but it was a list plot and had no nests.
#  RecObs[217,"FWSColonyID"]
iObs <- which(RecObs$FWSColonyID == "ME.0721")
RecObs[iObs, "PlotID"] <- "c4.16"
RecObs[iObs, "siteID"] <- "OOS"
RecObs[iObs, "Edit"] <- "Changed PlotID from c4.15 and siteID from L73."

# MCO: cd.1l (1 down and 5 up) is a ways off.  Not sure what to do for ME.0739.
# DDA:  This should be cd.1l
# MCO: cd.1l  is the listover.  I can fix that.
#  RecObs[221,"FWSColonyID"]
iObs <- which(RecObs$FWSColonyID == "ME.0739")
RecObs[iObs, "PlotID"] <- "cd.1l"
RecObs[iObs, "siteID"] <- "L79"
RecObs[iObs, "Edit"] <- "Changed PlotID from c4.16 and siteID from L03."

# New nests from talking to DDA 27 June 2024:
# OBJECTID, siteID, ColonyID
# 12 A27 848 	ce.1i? Co?
iObs <- grep("A27", RecObs$siteID)

# 115 A37 850
# 126 A18 851
# 137 A33 852
# OOS 846

```

## Spatial `RecObs`

```{r SpatialRecObs}
# Make a spatial file.  `tRecObs4` is a spatial file 
tRecObsSF2 <- tRecObs4 %>%
  filter(!is.na(Lng)) %>%
  mutate(
    Lng2 = Lng,
    Lat2 = Lat
  ) %>%
  st_as_sf(coords = c("Lng2", "Lat2"), crs = CRS.WGS84)

# This only had 137 observations, so it some locations are off.
tRecObsSF3 <- st_intersection(StrPlotSF %>%
                                select(Str_FK, PlotID_FK),
                              st_transform(x = tRecObsSF2, crs = st_crs(StrPlotSF))) %>%
  mutate(Str = ifelse(is.na(Str), Str_FK, Str),
         PlotID = ifelse(is.na(PlotID), PlotID_FK, PlotID))


RecObsSF <- st_as_sf(tRecObsSF2 %>%
                       filter(!is.na(Lng)),
                     coords = c("Lng", "Lat"),
                     crs = CRS.WGS84) %>%
  st_transform(crs = CRS.UTM19N)
if (lWrite) {
  st_write(RecObsSF, here::here("GIS/RecObsSF.kml"), delete_layer = TRUE)
}

rm(
   # RecObs2015,
   # RecObs2023,
   # tRecObs,
   # tRecObs2,
   tRecObs4,
   tRecObsSF2,
   tRecObsSF3)

```

## Check `RecObs`

Check stratum and plot for those that have it and add it where they are missing.

```{r CheckRecObs}
#| eval: false
tRecObsSF <- st_as_sf(RecObs, coords = c("Lng", "Lat"), crs = st_crs(4326))
tRecObsSF2 <- st_intersection(StrPlotSF %>%
                         select(tStr = Str, tPlotID = PlotID), 
                       st_transform(x = tRecObsSF, crs = st_crs(StrPlotSF))) %>%
  mutate(
    Str = ifelse(is.na(Str), tStr, Str),
    PlotID = ifelse(is.na(PlotID), tPlotID, PlotID)
  )

# Check of stratum and plot differences.  Resolve.
tRecObsSF2 %>%
  filter(tStr != Str | tPlotID != PlotID)

# Edit RecObs, to avoid the complications of dropping the spatial
# and rearranging the columns.

RecObs$Str <- ifelse(is.na(RecObs$Str), tRecObsSF2$Str, RecObs$Str)
RecObs$PlotID <- ifelse(is.na(RecObs$PlotID), tRecObsSF2$PlotID, RecObs$PlotID)
RecObs$FWSColonyID <- ifelse(
  grepl("\\.  \\.", RecObs$FWSColonyID),
  stringr::str_replace(RecObs$FWSColonyID, "  ", RecObs$Str),
  RecObs$FWSColonyID
)

```

## Ground Observations

Add the ground observations. Make a spatial file by adding the colony information.

```{r GroundObservations}
GroundObs2015 <- read_excel(path =                        "~/Profesional/GBHE.2015/Ori/GroundObs_DualFrame.xlsx", sheet = "Sheet1", skip = 1) %>%
  mutate(
    #      OBJECTID = row_number(),
    Year = 2015,
    Frame = "List",
    Svy = "GBHE",
    DatTim = as.character(Date),
    FWSColonyID = sprintf("ME.%04d", Colony),
    Obsr = "Gnd",
    Seat = "Dirt",
    Comment = NA
  ) %>%
  rename(Nest = `Total Nests`,
    Act = `Active Nests`) %>%
  select(Year,
         Frame,
         Svy,
         DatTim,
         FWSColonyID,
         Obsr,
         Seat,
         Nest,
         Act,
         Comment)

GroundObs2023 <- read_csv(file =              "~/Profesional/GBHE.2023/extdata/2023/GroundCounts.csv", col_types = "cciiii------cc") %>%
  mutate(
    #      OBJECTID = row_number(),
    Year = 2023,
    Frame = "List",
    Svy = "GBHE",
    DatTim = Date,
    FWSColonyID = sprintf("ME.%04d", Colony),
    Obsr = "Gnd",
    Seat = "Dirt",
    Nest = Total,
    Act = Active,
    Comment = paste(GN, Notes, sep = ";")
  ) %>%
  select(Year,
         Frame,
         Svy,
         DatTim,
         FWSColonyID,
         Obsr,
         Seat,
         Nest,
         Act,
         Comment)

tGroundObs <- bind_rows(GroundObs2015, GroundObs2023) %>%
  mutate(OBJECTID = row_number()) %>%
  select(OBJECTID,
         Year,
         Frame,
         Svy,
         DatTim,
         FWSColonyID,
         Obsr,
         Seat,
         Nest,
         Act,
         Comment)

tGroundObsSF <- left_join(tGroundObs,
                          AllColonySF %>%
                            select(FWSColonyID, Str, PlotID, Lng, Lat, geometry),
                          by = "FWSColonyID") %>%
  select(
    OBJECTID,
    Year,
    FWSColonyID,
    Frame,
    Svy,
    Str,
    PlotID,
    Lng,
    Lat,
    DatTim,
    Obsr,
    Seat,
    Nest,
    Act,
    Comment,
    geometry
  )

# These 21 ground observations are not in the colony list I have
tGroundObsSF %>%
  filter(is.na(Lng))

GroundObsSF <-     tGroundObsSF %>%
  filter(!is.na(Lng)) %>%
  st_sf()

if (lWrite) {
  st_write(GroundObsSF,
           here::here("GIS/GroundObsSF.kml"),
           delete_layer = TRUE)
}
rm(GroundObs2015, GroundObs2023, tGroundObs, tGroundObsSF)

```
## Non-Spatial Ground Observations


```{r GroundObsNonSpatial}
GroundObs <- st_drop_geometry(GroundObsSF)
if (lWrite) {
  readr::write_csv(GroundObs, here::here("data/GroundObs.csv"), na = "")
}

```
## Save Data

Save all the tables we need for the estimation. The spatial data are already in `gis` and the non-spatial tables in `data`, so they can be accessed easily by `locdata` which is more general than `base::data`. `Load`ing the save file makes data access easy as if this were an `R` project.

```{r SaveAll}
save(
  CRS.WGS84,
  CRS.Alpers,
  CRS.UTM19N,
  cYear,
  nYear,
  cFrame,
  nFrame,
  cDFStr,
  nStr,
  cYrStr,
  nYrStr,
  cFS,
  nFS,
  MEBoundarySF,
  StrSF,
  ColonySF,
  Colony,
  StrPlotSF,
  List,
  Plot,
  GRTSSmp,
  tblFlightInfo,
  tblPlotDetails,
  tblObservations,
  RecObs,
  NewColonies,
  GroundObsSF,
  GroundObs,
  file = here::here("output/GBHE.RData")
)

```
